A Middleware for Symmetric Peer-to-Peer Applications

### 2023-03-12 tpds-23/tpds-23.tex
...

A common approach towards symmetric applications is to introduce a middleware
to orchestrate events and time in the network~\cite{gals,croquet}.
This way, applications developers can rely on middleware primitives to trigger
events, which are intercepted and synchronized in a global shared timeline
across the network.
Developers must also restrict themselves to deterministic and stateless calls
only, such that execution can be equally reproduced in all nodes according to
the shared timeline. %~\cite{TODO:det_reprod}.
However, current solutions depend on a central server to interconnect network
nodes and determine the shared timeline.
Therefore, applications may experience failures, unless users are permanently
connected, with no partitions in the network, particularly from the server.
In contrast, a decentralized solution could be resilient to partitions, and
ultimately, even work offline as promoted by local-first software~\cite{local}.

In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications}, which do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer
networks~\cite{p2p.survey}.
%
Events are flooded in the graph and are applied with a small deadline delay to
compensate the network latency.
To deal with events received out of order or too late, a time machine can
rollback a peer to a previous state and reapply the events in order, such that
a single consistent timeline is established across the network.
%
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.
%
Our middleware provides \emph{perceptive consistency}~\cite{melding}, which
ensures that an event triggered by any peer at time $t$ is perceived by all
peers simultaneously at time $t+\Delta$.
%
As the main limitation, the middleware cannot ensure \emph{instantaneity} for
events in most scenarios, which would require $\Delta$ to be less than a
perceptive threshold~\cite{melding}.
We also that assume peers are non malicious in the sense that they do not lie
about the time in which events occurred locally.

@@

Croquet~\cite{croquet,croquet.site} guarantees bit-identical real-time behavior
for users in collaborative distributed environments.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
fixed periodic ticks (e.g., every $50ms$), such that all nodes in the network
advance together according to a synchronized global shared clock.
If a user generates an event, it is sent back to the reflector, which
broadcasts the event in the next tick, which all nodes apply in sync.
%
This mechanism is similar to \emph{bucket synchronization}~\cite{gautier}, in
which events are assigned to time buckets to be applied later.
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to request the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network (a), in which nodes advance in sync (b), and in which event broadcasts
and outcomes depend on the central server to be online (c).

@@

@@

The performance of Croquet~\cite{croquet} depends exclusively on the network
latency, which restricts the application to a maximum FPS.
For instance, since all frame ticks need to be transmitted over the network,
a latency of 50ms supports at most 20 FPS.
Also, the higher is the FPS, the higher is the network traffic and server load,
which may also degrade performance.
%
GALS~\cite{gals} detaches latency from FPS, since each node has its own
timeline.
However, the protocol requires two round trips to the server, which degrades
performance considerably.
The authors claim to achieve a responsiveness of 350ms in applications with
25 nodes at 25 FPS, where 350ms represents the time it takes for all nodes to
apply an event since its initial occurrence.

@@

We propose a middleware to program \emph{symmetric peer-to-peer applications}
in which interacting peers conform to identical behavior without a centralized
coordination server.
%
The middleware API shares similar limitations with centralized
alternatives~\cite{gals,croquet}, being restricted to deterministic and
stateless calls. %, and only supports pre-allocated memory.
%
%To our knowledge, it is the first middleware to support bit-identical real-time
%behavior in decentralized peer-to-peer networks.
%
Our main contribution is a time machine that supports rollbacks based on memory
snapshots and deterministic simulation.
This allows to decentralize the network timeline, since conflicting peers can
resynchronize their state through time travels.
%
%We assume that peers are non-malicious and that they can assign themselves
%unique identifiers.
%
The middleware provides \emph{perceptive consistency}~\cite{melding}, which
adds a time deadline to reach consistency, ensuring that an event triggered by
any peer at time $t$ is perceived by all peers simultaneously at time
$t+\Delta$.

>>>>

A common approach towards symmetric applications is to introduce a middleware
to orchestrate events and time in the network~\cite{gals,croquet2}.
This way, applications developers can rely on middleware primitives to trigger
events, which are intercepted and synchronized in a global shared timeline
across the network.
Developers must also restrict themselves to deterministic and stateless calls
only, such that execution can be equally reproduced in all nodes according to
the shared timeline. %~\cite{TODO:det_reprod}.
However, current solutions depend on a central server to interconnect network
nodes and determine the shared timeline.
Therefore, applications may experience failures, unless users are permanently
connected, with no partitions in the network, particularly from the server.
In contrast, a decentralized solution could be resilient to partitions, and
ultimately, even work offline as promoted by local-first software~\cite{local}.

In this work, we propose a middleware for \emph{symmetric peer-to-peer
applications}, which do not rely on central servers for coordination.
Peers in the application form a dynamic network graph and communicate only
with direct neighbours, as in typical unstructured peer-to-peer
networks~\cite{p2p.survey}.
%
Events are flooded in the graph and are applied with a small deadline delay to
compensate the network latency.
To deal with events received out of order or too late, a time machine can
rollback a peer to a previous state and reapply the events in order, such that
a single consistent timeline is established across the network.
%
Our main contribution is the design of a middleware to ensure that all peers
    (a) meet event deadlines,
    (b) advance in sync in real time, and
    (c) can leave and join the network and remain symmetric.
%
Our middleware provides \emph{perceptive consistency}~\cite{melding}, which
ensures that an event triggered by any peer at time $t$ is perceived by all
peers simultaneously at time $t+\Delta$.
%
As the main limitation, the middleware cannot ensure \emph{instantaneity} for
events in most scenarios, which would require $\Delta$ to be less than a
user-observable threshold~\cite{melding}.
We also that assume peers are non malicious in the sense that they do not lie
about the time in which events occurred locally.

@@

Croquet~\cite{croquet,croquet2,croquet.site} guarantees bit-identical real-time
behavior for users in collaborative distributed environments.
%
As detailed in Figure~\ref{fig.algos}, a centralized \emph{reflector} issues
fixed periodic ticks (e.g., every $50ms$), such that all nodes in the network
advance together according to a synchronized global shared clock.
If a user generates an event, it is sent back to the reflector, which
broadcasts the event in the next tick, which all nodes apply in sync.
%
This mechanism is similar to \emph{bucket synchronization}~\cite{gautier}, in
which events are assigned to time buckets to be applied later.
%
Croquet takes periodic snapshots of the whole application state in order to
support late joins to a running session.
The new node just needs to request the latest snapshot and simulate the
remaining events to reach the current running state.
%
As indicated in Figure~\ref{fig.table}, Croquet relies on a centralized
network (a), in which nodes advance in sync (b), and in which event broadcasts
and outcomes depend on the central server to be online (c).

@@

The main advantage of CRDTs is that they can be designed in such a way that
concurrent modifications to shared state can preserve intent~\cite{peritext},
regardless of the order in which they are applied.
%
This is not the case for the symmetric application middlewares described in the
previous section.
Nevertheless, programmers can still use CRDTs on top of a middleware, since
they are nothing more than data structures and merging algorithms.

@@

The performance of Croquet~\cite{croquet2} depends exclusively on the network
latency, which restricts the application to a maximum FPS.
For instance, since all frame ticks need to be transmitted over the network,
a latency of 50ms supports at most 20 FPS.
Also, the higher is the FPS, the higher is the network traffic and server load,
which may also degrade performance.
%
GALS~\cite{gals} detaches latency from FPS, since each node has its own
timeline.
However, the protocol requires two round trips to the server, which degrades
performance considerably.
The authors claim to achieve a responsiveness of 350ms in applications with
25 nodes at 25 FPS, where 350ms represents the time it takes for all nodes to
apply an event since its initial occurrence.

@@

We propose a middleware to program \emph{symmetric peer-to-peer applications}
in which interacting peers conform to identical behavior without a centralized
coordination server.
%
The middleware API shares similar limitations with centralized
alternatives~\cite{gals,croquet2}, being restricted to deterministic and
stateless calls. %, and only supports pre-allocated memory.
%
%To our knowledge, it is the first middleware to support bit-identical real-time
%behavior in decentralized peer-to-peer networks.
%
Our main contribution is a time machine that supports rollbacks based on memory
snapshots and deterministic simulation.
This allows to decentralize the network timeline, since conflicting peers can
resynchronize their state through time travels.
%
%We assume that peers are non-malicious and that they can assign themselves
%unique identifiers.
%
The middleware provides \emph{perceptive consistency}~\cite{melding}, which
adds a time deadline to reach consistency, ensuring that an event triggered by
any peer at time $t$ is perceived by all peers simultaneously at time
$t+\Delta$.
